### 2018-03-16 paper.Rmd
added acknowledgements stub
intro - closed #2

  Enter author note here.

@@

Children's language acquisition is a core puzzle in cognitive science. How are the representations underlying the generative use of language learned over their first years of life? Because a language is comprised of many individual pieces of information -- the names of particular objects, the marker for the past tense, the position of determiners -- studying language development is fundamentally about studying learning from environmental input, which is the only source of information on these specifics. Further, children are often reluctant experimental participants; often their most revealing and most sophisticated uses of language emerge in naturalistic play. Thus, language acquisition research must be grounded in the ecological behavior of children, both their language input and their language production in context.

@@

Established in 1984 to address this issue, the Child Language Data Exchange System (CHILDES) aims to make transcripts and recording of children's language available for researchers via a free resource [@macwhinney2000]. CHILDES archives tens of thousands of transcripts of children’s speech, making it an critical resource for characterizing children’s early productive language use as well as their language environment. Researchers have used the dataset for studies including estimating children’s latent vocabulary size, characterizing their lexical diversity, evaluating the prevalence of negative examples to children, and retrieving the frequencies of various syntactic constructions, among thousands of others. As the first major effort to consolidate and share transcripts of child language, CHILDES has been a pioneer in the move to curate and disseminate large-scale behavioral datasets publicly.

Yet there are some challenges working with CHILDES, both for students and for advanced. For beginners with fewer technical tools, mastering the system can be tricky. The CHILDES ecosystem uses a special file format (CHAT), which is read by a command-line tool (CLAN) with a host of specific arguments. Mastering the CHAT codes and the CLAN tool can be daunting to navigate for a beginning researcher or student, decreasing the ease of integrating hands-on exploration of the dataset into classes.

On the opposite end of the spectrum. for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. Many researchers who would like to ingest the entire corpus (or some large subset) into a computational analysis must write one-off parsers of the CHAT format to extract the subset of the data they would like to use [EXAMPLES]. This practice is problematic for a number of reasons. Not only is effort wasted in rewriting the same parser again and again, but this process introduces a possible source of error in data handling. Further, these scripta are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017]. Finally, the CHILDES corpus itself is a moving target; thus computational work using the entire corpus at one time point may include a different set of data (due to new contributions or revisions).

Our goal in the current manuscript is to describe a system for extending the functionality of CHILDES to address these issue. childes-db is a database-formatted copy of the CHILDES data that allows programmatic access through an application programming interface (API). This infrastructure allows the creation of web applications for browsing and visualizing the data easily, facilitating classroom use of the dataset. Further, it can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format. The database is versioned for access to previous images, allowing reproducibility of particular analyses.

<!-- Uses -->
<!-- Uptake -->
<!-- Intake -->
<!-- Theory / growing interest in child language growth (roger brown etc) -->
<!-- Why dataset is successful in light of theoretical interests discussed -->
<!-- Challenges working with CHILDES -->
<!-- Expanding the range of use cases (beginner / expert) -->
<!-- Pulling out for modeling:  -->
<!-- People write their own parsers/pipelines -->
<!-- Duplicated effort -->
<!-- Versioning and reproducibility -->
<!-- Can’t easily integrate clan calls with pipeline -->
<!-- Solution (childes-db) -->

<!-- At its core is a series of linked tabular data stores, currently implemented in `MySQL`, which can be accessed from dozens of programming languages. The second is an `R` package called `childesr` that provides abstractions for the most common complex queries, obviating the need to write `SQL`. `childesr` exposes preprocessed, tabular data from CHILDES through its application programming interface (API), providing researchers with the opportunity to exploit conveniently formatted data in a familiar programming environment. Finally, we provide  a collection of browser-based interactive applications that enable visual exploration of CHILDES, allowing researchers to easily view and download useful information from the corpus including mean length of utterance (MLU) and word frequency counts. -->

@@

High level discussion of features

## Childes-db

@@

Technical Implementation

@@

Versioning

@@

```{r}
stats <- get_speaker_statistics(role = "Target_Child")

>>>>

  This work supported by a Jacobs Foundation Fellowship to MCF.

@@

Children's language acquisition is a core puzzle in cognitive science. How are the representations underlying the generative use of language learned over children's first years? Because a language is comprised of many individual pieces of information -- the names of particular objects, the marker for the past tense, the position of determiners -- language development is fundamentally a question of learning from environmental input, which is the only source of information on these specifics. Further, children are often reluctant experimental participants; often their most revealing and most sophisticated uses of language emerge in naturalistic play. Thus, language acquisition research must be grounded in the ecological behavior of children, both their language input and their language production in context.

@@

Established in 1984 to address this issue, the Child Language Data Exchange System (CHILDES) aims to make transcripts and recording of children's language available for researchers via a free resource [@macwhinney2000]. CHILDES archives tens of thousands of transcripts of children’s speech, making it an critical resource for characterizing children’s early productive language use as well as their language environment. As the first major effort to consolidate and share transcripts of child language, CHILDES has been a pioneer in the move to curate and disseminate large-scale behavioral datasets publicly.

Since its inception, a tremendous corpus of research has made use of CHILDES data. Individual studies are too numerous to list, but  classics include studies of morphological over-regularization [@marcus1992], distributional learning [@redington1998], word segmentation [@goldwater2009], the role of frequency in word learning [@goodman2008] and many others. Some studies analyze individual examples in depth [e.g., @snider2007] while others work with the corpus as a whole [e.g., @meylan2017]. These are just a few of the thousands of studies that take CHILDES corpora -- transcripts, annotations, and associated media -- as starting points.

Yet there are some challenges working with CHILDES, both for students and for advanced users. For beginners with a lower level of technical experience, mastering the system can be tricky. The CHILDES ecosystem uses a special file format (CHAT), which is stored in a plain text format but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological or morphological structure of the utterance. These files are analyzed via by a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU). While this system is flexible and powerful, mastering the CHAT codes and the CLAN tool can be daunting to navigate for a beginner or student. These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- can be tricky to build into a processing pipeline written in a modern analytic platform such as python or R. Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analysis typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998,@yang2013,@meylan2017].

This practice is problematic for a number of reasons. Effort wasted in rewriting the same parser again and again. In addition, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard. Further, these scripta are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017]. This lack of computational reproducibility is widely recognized as problematic because it does not allow researchers to verify or build on published researcher [@donohoe2010,@stodden2016]. Finally, the CHILDES corpus itself is a moving target;  computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to new contributions or revisions).

In the current manuscript, we describe a system for extending the functionality of CHILDES to address these issues. Our systemn, childes-db, is a database-formatted mirror of CHILDES that allows programmatic access through an application programming interface (API). This infrastructure allows the creation of web applications for browsing and visualizing the data easily, facilitating classroom use of the dataset. Further, it can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format. The database is versioned for access to previous images, allowing computational reproducibility of particular analyses.

We begin by describing the architecture of childes-db and the web applications that we provide. We next describe the `childesr` API, which allows for programmatic access to the data. We then present two worked examples of specific uses of the system using both the web app and the API. The first examines the frequency of color words in children's input; the second quantifies the growth of lexical diversity across genders. We end by presenting some suggestions for the use of the childes-db web frontend in classroom teaching.

@@

We first present the design and technical approach of childes-db, then describe the interactive analysis tools provided by our web-based frontend.

## Design and technical approach

@@

### Technical Implementation

@@

### Versioning

@@

```{r warning=FALSE}
stats <- get_speaker_statistics(role = "Target_Child")
